---
title: 'blogdown updates prompted a website overhaul: These are my notes'
author: A. Solomon Kurz
date: '2021-04-27'
draft: false
excerpt: "The purpose of this post is to highlight some of the steps I took to rebuild my academic-style blogdown website. At a minimum, I'm hoping this post will help me better understand how to set up my website the next time it needs an overhaul. Perhaps it will be of some help to you, too."
layout: single
tags:
- blogdown
- hugo
- Netlify
- R
- tutorial
lastmod: '2021-04-26T11:27:11-07:00'
featured: no
bibliography: /Users/solomonkurz/Dropbox/blogdown/content/post/my_blog.bib
biblio-style: apalike
csl: /Users/solomonkurz/Dropbox/blogdown/content/post/apa.csl  
link-citations: yes
---

```{r, echo = F, cache = F}
options(width = 120)
```

## Purpose

A few weeks ago, I was preparing to release the second blog post in a two-part series (you can find that post [here](https://solomonkurz.netlify.app/post/2021-04-22-effect-sizes-for-experimental-trials-analyzed-with-multilevel-growth-models-two-of-two/)). During the editing process, I had rendered the files into HTML and tried posting the draft to my website. Everything looked fine except that the figures wouldn't render. I hadn't seen this behavior before and I figured it had to do with some software update. When I checked, the [**blogdown** package](https://CRAN.R-project.org/package=blogdown) [@R-blogdown; @xieBlogdown2017] had indeed recently updated. I'd also noticed the great [Alison Hill](https://alison.rbind.io) had recently posted a few blogs on **blogdown**-related topics, so I figured it was time for a refresh. 

The purpose of this post is to highlight some of the steps I took to rebuild my website and recover my figure-rendering game. At a minimum, I'm hoping this post will help me better understand how to set up my website the next time it needs an overhaul. Perhaps it will be of some help to you, too.

### I don't cover everything.

This post is not an exhaustive introduction to **blogdown**. For that, you have the ebook by Xie, Hill, and Thomas [-@xieBlogdown2017], [*blogdown: Creating websites with R markdown*](https://bookdown.org/yihui/blogdown/). A difficulty with that book is the authors designed it to cover a broad range of applications[^1], which means there isn't enough room to cover every special case. A further difficulty is **blogdown** is something of a high-level interface for an array of more-specific software solutions (e.g., [Netlify](https://www.netlify.com/), [Hugo](https://gohugo.io/)), each of which has its own quirks. I am not going to introduce these in any great detail, either. For those purposes, you have the relevant reference manuals and other documentations available on the web.

### This isn't the only way.

There are any number of ways one could make an academic website via **blogdown**. Hill provided one workflow in her post, [*Up & running with blogdown in 2021*](https://alison.rbind.io/post/new-year-new-blogdown/), upon which I'll be drawing heavily throughout this post. Some of my steps will follow a different order from hers, based on what seemed right, for me.

## Foundations

Hill organized her aforementioned blog post, [*Up & running with blogdown in 2021*](https://alison.rbind.io/post/new-year-new-blogdown/), as if one were building their **blogdown** website from scratch. After futzing around with different strategies, I recommend this approach even if you've had a **blogdown** website up and running for a while. If you haven't updated your website recently, archive your old files and build the new one from the ground up. Here's the first step:

### Step 1. GitHub.

Log on to your GitHub account and start a fresh repo. Name it something website-y. I named mine "blogdown", which you can find at [https://github.com/ASKurz/blogdown/](https://github.com/ASKurz/blogdown/). If you need a refresher on GitHub, let the great Jenny Bryan lead you, [here](http://happygitwithr.com/) [@bryanHappyGitGitHub2020]. 

### Step 2. RStudio projects.

Make a fresh RStudio project[^2] to go along with your fresh GitHub repo. Within RStudio, you can do this by clicking through `File > New Project > Version Control > Git`. Next, you'll want to paste in the URL from your GitHub repo. If you haven't done something like this, before, go back online to your repo and click the green button near the top that's labeled "Clone or download." A URL will appear in there. That's what you'll be pasting into your new RStudio project, which will connect it to your GitHub repo. Hill discussed this [here](https://alison.rbind.io/post/new-year-new-blogdown/#step-1-create-repo).

## **blogdown** mini launch

### Step 3. Make a default **blogdown** site.

If you haven't already, install the current version of **blogdown** by executing `install.packages("blogdown")`. Restart **R**, if necessary. Now within a fresh session within your RStudio project, execute `blogdown::new_site(theme = "wowchemy/starter-academic")`[^3]. Over the next minute or two, you'll see a handful of new files pop up in your project folder. In your console, you'll probably notice the prompt: "Want to serve and preview the site now? (y/n)". I recommend executing `y`, which will return a preview of your default **blogdown** website in the RStudio Viewer panel.

### Step 4. `.gitignore`.

I don't know that you have to do this right now, but a good early step is to make a `.gitignore` file. Following Hill ([here](https://alison.rbind.io/post/new-year-new-blogdown/#using-github)), you can do this by executing `file.edit(".gitignore")`. Then go ahead and enter this content to the file:

```{r, eval = F}
.Rproj.user
.Rhistory
.RData
.Ruserdata
.DS_Store
Thumbs.db 
/public/
/resources/
```

Once you save the changes, you might execute `blogdown::check_gitignore()` to confirm you're okay.

If you haven't done so, yet, this would be a good time to commit your changes and push them to GitHub. Again, if this is new to you, let @bryanHappyGitGitHub2020 lead you.

## Netlify

### Step 5. Sign up and deploy.

Hill recommended you both build and host your **blogdown** website on Neflify (see [here](https://alison.rbind.io/post/new-year-new-blogdown/#step-1-create-repo)). I'm not going to argue. Go to [https://www.netlify.com/](https://www.netlify.com/) and either log in or sign up using your GitHub account. Even if you already have a Netlify account, I recommend making a new Netlify site with `New site from Git > Continuous Deployment: GitHub`. You'll then need to select your fresh GitHub repo, from above, to connect it to Netlify. This may require you to follow the prompts to actually navigate to GitHub, enable the connection, there, and then follow back to Netlify. Once you're back in Netlify, leave settings at their defaults and select `Deploy Site`. You should be directed to a page with a header called *Production deploys* somewhere in the middle of the screen. After a minute of two, Netlify will finish deploying your site. 

### Step 6. Customize your Netlify subdomain.

When you create your new site, Netlify will have automatically generated a subdomain name following the form `random-word-12345`. You should be able to see this at the top of your screen. This subdomain name will be part of your web address. You're at liberty to keep the default name, if you want. But you can customize your subdomain name by navigating to `Site settings > General > Site details`. Then click the gray button named `Change site name`. In the field, I renamed my subdomain to `solomonkurz`. As a result, my website is deployed at [https://solomonkurz.netlify.app/](https://solomonkurz.netlify.app/). Once you save this change, your website should be available at your customized address almost instantly.

## Start customizing

### Step 7. `config.yaml`.

You should see a `config.yaml` file in the root directory of your RStudio project folder. Open it and update the info at the top. Here's what I changed:

```{r, eval = F}
title: A. Solomon Kurz # Website name
baseurl: 'https://solomonkurz.netlify.app/'
copyright: '© A. Solomon Kurz (2021)' 
```

### Step 8. `netlify.toml`.

Next, check for a `netlify.toml` file in your root directory. You can also open or create it by executing `blogdown::config_netlify()`. Once the file's open, make sure Hugo is installed and matched up by executing `blogdown::check_hugo()`. I needed to update my `netlify.toml` file to include the following:

```{r, eval = F}
[build.environment]
  HUGO_VERSION = "0.82.0"
```

If this seems scary, Hill discussed it in greater detail [here](https://alison.rbind.io/post/new-year-new-blogdown/#step-5-publish-site).

### Step 9. `.Rprofile`.

Still following Hill ([here](https://alison.rbind.io/post/new-year-new-blogdown/#step-4-create-content)), make an `.Rprofile` file by executing `blogdown::config_Rprofile()`. Then customize some of the settings within `options()`, as desired. Mine now read:

```{r, eval = F}
options(
  # to automatically serve the site on RStudio startup, set this option to TRUE
  blogdown.serve_site.startup = FALSE,
  # to disable knitting Rmd files on save, set this option to FALSE
  blogdown.knit.on_save = TRUE,
  # build .Rmd to .html (via Pandoc); to build to Markdown, set this option to 'markdown'
  blogdown.method = 'html',
  # fix Hugo version
  blogdown.hugo.version = "0.82.0",
  # These changes are based on Alison Hill's posts:
  # https://alison.rbind.io/post/2019-02-21-hugo-page-bundles/#project-specific-rprofile
  # and 
  # https://alison.rbind.io/post/new-year-new-blogdown/#step-4-create-content
  blogdown.author = "A. Solomon Kurz",
  blogdown.ext = ".Rmd",
  blogdown.subdir = "post",
  blogdown.yaml.empty = TRUE,
  blogdown.new_bundle = TRUE
)
```

In her post, Hill stressed it's important to restart your RStudio session after you save any changes to `.Rprofile`. If you haven't done so, in a while, this is probably a good time to commit and push your files to GitHub.

## Content

